%!TEX root = ../report.tex

\chapter{Conclusions}
\label{chap:chap5}

\section*{}


\section{Summary} % (fold)
\label{sec:summary}

  The proposed thesis focused on delivering an enhanced user experience when discovering new music in Spotify's Environment.
  By using RAMA's concept applied in the developed prototype, the user experience when discovering new music as been greatly increased.
  The users felt that RAMA's Spotify Application was natural and intuitive.

  The amount of services that use visual tools for recommending music to users are not that many, although the ones shown here (\ref{chap:chap2}) are not representative of the whole spectrum.

  Given the overview of the possibilities, creating a Spotify Application to apply RAMA's concept proved to be the best option to take.

  The proposed modules \ref{intro:obj4} and \ref{item:obj5} were implemented.
  The developed prototype proved to work after the beta-testing results.
  Although, there are a lot of improvements to do, the final result was very appealing to the users.
  All of the beta-testers liked the visual experience and the majority responded positively about using the application in a regular basis to discover new music.

  All of the developed material (code, documentation, wiki, screenshots, demos) can be found in the project's code repository (\url{http://github.com/carsy/rama-spotify}) as well as in the appendix to this document.

  % NOTE
  % articles to be submitted

% section summary (end)

\section{Discussion} % (fold)
\label{sec:discussion}

  By introducing a visualization tool into a complete service like Spotify, the users felt that their experience with RAMA's application improved their abilities to find new music.
  The tests' results show that RAMA's Spotify Application is a successful approach to music discovery and recommendation.
  Although, the final results point in that direction, after the experiments, 3 beta-testers stated that their music listening habits are not focused on the music artists they are listening to. Instead, they simply pay attention to the songs (mostly, the popular ones), and so, their playlists are track-driven, not artist-driven.
  That might have had presented a problem to those users, since the focus of RAMA is the relations between the artists.
  However, Spotify's API's recommendation system proved to please those users, who started to pay more attention to the name of the artists they listen to.

  Services like Spotify or Rdio, offer a complete set of features that range between playing every track on their catalogue, to saving albums for offline mobile listening.
  With such a vast music catalogue, the user might feel lost and not very motivated to find new music.
  Although these services continue to add features like Spotify's "Radio"~\cite{spradio}, the user finds it hard to compute such a large world of available music.

  By introducing this visual aid to music artist's relations, RAMA succeeds in letting the user understand and explore better the whole spectrum of available music.

% section discussion (end)

\section{Future Work} % (fold)
\label{sec:future_work}

  From the developed work so far, the need for a more segmented testing might have been called for, in order to improve the results obtained, as well to better understand the user's needs.

  During the prototype's development and the beta-testing experiments, there were several additional features suggested for RAMA's Spotify Application.
  Most of them were not included in the prototype because of the limited time frame for development, while others seemed to stray from RAMA's focus.
  Nonetheless, every idea is important and might someday contribute for RAMA's future features.

  \paragraph*{Place the actions buttons (expand, new map and delete) near the artist node} \hfill \\
  \indent Some users complained that it was hard to notice the actions buttons (one user did not even notice them).
  Maybe placing the buttons near the artist node and only show them while hovering it would make more sense.

  \paragraph*{Make it clear that the tags are clickable to the user} \hfill \\
  \indent Some users noticed that it was not apparent at first that the upper tags were clickable, they thought it was just an extra information, not something they could interact with.

  \paragraph*{Allow to click-and-hold to preview listening to an artist} \hfill \\
  \indent This Spotify feature could be applied to the nodes so that the user could browse the graph much faster.

  \paragraph*{Select multiple artist or tags} \hfill \\
  \indent Allow the user to select multiple artists (or tags), in order to, for example, generate playlists from that selection.

  \paragraph*{Let the graph update automatically with the current playing artist} \hfill \\
  \indent In order to keep RAMA more alive, the graph would refresh automatically when a new artist starts playing, with the new updated artist as root.
  Maybe an option to keep the graph "locked" would allow the user choose if he/she wants that behaviour or not.

  \paragraph*{Improve artists recommendation} \hfill \\
  \indent The assumption that Spotify's recommendations are valid might not be so desirable.
  Although the results are satisfactory, without knowing the reasons behind those choices of connections, the artists recommendations become inflexible.
  To improve this, one could fully use Echonest's API to fetch the similar artists~\cite{echonestsimilar}.
  The amount of query personalization is much more attractive than Spotify's.

  \paragraph*{Improve the connection between the artists and the tags} \hfill \\
  \indent As an improvement of the previous idea, the connections between the artists and the overlaying tags could become more related, and so, the intersection of tags between artists could be used to improve the recommendations (creation of connections). 

  \paragraph*{Include an alternate view with tags as nodes, instead of artists} \hfill \\
  \indent Some users suggested that the tags overlay was an interesting point of view, and so, if the graph could be drawn using the tags instead, the resulting map could be an alternative view of the application.

  \paragraph*{Cluster groups of artists} \hfill \\
  \indent Cluster groups of artists to create a geographical representation of what the entire network of available artists could be like.

  \paragraph*{Tag mind map} \hfill \\
  \indent In a way to try and complement the two previous ideas, a clustering of the tags would create an interesting mind map of the current genres.
  And since there are indeed music genres that are similar (like metal and celtic metal) the map would be browsable.

  \paragraph*{Information} \hfill \\
  \indent Add an information UI component to help the users understand better the visualization parameters (a tutorial-like component) - although those parameters are advanced settings (not strictly necessary for the user to interact with) it might be a good idea to give more power to the user and make these settings easy to understand.

% section future_work (end)